---
title: 'Reproducible Research: Peer Assessment 1'
output: html_document
keep_md: yes
pdf_document: default
---

```{r global_options, include=FALSE}
options(scipen = 999, digits = 2)
```

## Loading and preprocessing the data
Load the data and make the date a date type.
```{r}
rm(list=ls(all=TRUE))
actdat <- read.csv("activity.csv", na.strings = "NA", header = TRUE)
actdat$date <- as.Date(actdat$date)
```

## What is mean total number of steps taken per day?

```{r}
TotalSteps <- aggregate(steps~date, data=actdat, sum, na.rm=TRUE)
hist(TotalSteps$steps, breaks=10)
meansteps <- mean(TotalSteps$steps)
mediansteps <- median(TotalSteps$steps)
```
The mean steps is `r meansteps`.  
The median steps is `r mediansteps`.  

## What is the average daily activity pattern?
```{r}
DayInterval<-aggregate(steps~interval,data=actdat,mean,na.rm=TRUE)
plot(steps~interval,data=DayInterval,type="l")
theDay <- DayInterval[which.max(DayInterval$steps),]$interval
```
The interval that has the most steps is `r theDay`.  

## Imputing missing values
I will assume missing intervals are 0 steps.  There is no real good way to do it other than this since we don't know if the device is malfuncitoning or just off and if the person is really walking.

```{r}
missing <- sum(is.na(actdat$steps))

actFill <- actdat
actFill[is.na(actFill$steps), "steps"] <- 0

TotalSteps <- aggregate(steps~date, data=actFill, sum, na.rm=TRUE)
hist(TotalSteps$steps, breaks=10)
fillMean <- mean(TotalSteps$steps)
fillMedian <- median(TotalSteps$steps)
```

Total missing values is `r missing`.  
The filled in mean is `r fillMean` and the non-filled is is `r meansteps`.  
The filled in median is `r fillMedian` and the non-filled is is `r mediansteps`.  
Filling in the data alters the mean and median by adding vaules that move the mean and median.  

## Are there differences in activity patterns between weekdays and weekends?

```{r}
daytype = function(x)
			{
				if(x %in% c('Saturday', 'Sunday')){
						return('Weekend')
				}

				return('Weekday')
			}
		
actFill$dayname = weekdays(actFill$date)	
actFill$daytype <- as.factor(apply(as.array(actFill$dayname), 1, FUN=daytype))

avgDayType = aggregate(steps~interval+daytype, actFill ,mean)

library(lattice)
xyplot(steps~interval|factor(daytype),data=avgDayType, aspect=1/2,type="l")
```			